True versus apparent shapes of bow shocks

### 2017-12-06 papers/Papers1234/_region_.tex
Submitted first paper! To MNRAS and arxiv

\section{Analytic derivation of thin-shell bow shape parameters}
\label{sec:thin-shell-shapes}

In this appendix, we provide analytic calculations of the planitude,
alatude, and asymptotic opening angle for the wilkinoid, cantoids, and
ancantoids.  We first consider the most general case of the
ancantoids, and then show how results for cantoids and the wilkinoid
follow as special cases.

@@

To find the alatude, \(\Lambda = R_{90} / R_0\), we use
equation~\eqref{eq:crw-angles} at \(\theta = \ang{90}\) to write
\begin{equation}
  \label{eq:Lambda-from-theta-1-90}
  \Lambda = \frac{D} {R_0} \tan \theta_{1,90} \ , 
\end{equation}
where \(\theta_{1,90} = \theta_1(\theta = \ang{90})\), which, following
equation~\eqref{eq:ancantoid-theta-theta1-implicit}, must satisfy
\begin{equation}
  \label{eq:theta-1-90-implicit}
  \theta_{1,90} \cot \theta_{1,90}  = 1 - \frac{2 \beta}{k + 2} \ . 
\end{equation}
Combining \eqref{eq:Lambda-from-theta-1-90} and
\eqref{eq:theta-1-90-implicit} with \eqref{eq:again-R0-over-D} yields
\begin{equation}
  \label{eq:Lambda-beta-xi-theta-1-90}
  \Lambda = \frac{ \left(1 + \beta^{1/2}\right) \,\theta_{1,90}} {\beta^{1/2} \left(1 - \xi_k \beta\right)} \ ,
\end{equation}
where
\begin{equation}
  \label{eq:xi-k}
  \xi_k = \frac{2} {k +2} \ .
\end{equation}
We now take the Taylor expansion of equation~\eqref{eq:theta-1-90-implicit} to find
\begin{equation}
  \label{eq:theta-1-90-Taylor}
  \theta_{1,90}^2 + \tfrac{1}{15}  \theta_{1,90}^4 + \mathcal{O}(\theta_{1,90}^6)
  = 3 \xi_k \beta \ , 
\end{equation}
which, if \(\theta_{1,90}\) is small, has the approximate solution
\begin{equation}
  \label{eq:theta-1-90-approx}
  \theta_{1,90} \approx \left( \frac{3 \xi_k \beta} {1 + \tfrac15 \xi_k \beta} \right) \ .
\end{equation}
Substituting back into equation~\eqref{eq:Lambda-from-theta-1-90}
yields an approximate value for the alatude of
\begin{equation}
  \label{eq:Lambda-approx}
  \text{ancantoid} \quad
  \Lambda \approx \frac {(3 \xi_k)^{1/2} \left( 1 + \beta^{1/2} \right)}
  { \left( 1 + \tfrac15 \xi_k \beta \right)^{1/2} \left( 1 - \xi_k \beta \right)} \ .
\end{equation}
This approximation is surprisingly accurate, with a relative error of
order \(1\%\) even for \(\beta\) as large as \(0.5\) with \(k = 0\).

@@

Since \(\Pi\) and \(\Lambda\) depend on only that portion of the inner wind
emitted in the forward hemisphere, \(\theta \le \ang{90}\), the results for the
cantoids can be found by taking \(k = 0\), in which case
equations~(\ref{eq:C-k-beta}, \ref{eq:final-planitude}, \ref{eq:xi-k},
\ref{eq:Lambda-approx}) yield
\begin{gather}
  \label{eq:cantoid-Pi-Lambda}
  \text{cantoid} \quad
  \begin{cases}
    \quad \Pi &= \dfrac {5} {3 \left( 1 - \beta^{1/2} \right)}\\[12pt]
    \quad \Lambda &= \dfrac { \sqrt 3} {\left( 1 + \tfrac15 \beta \right)^{1/2} \left( 1 - \beta^{1/2} \right)} \ .
  \end{cases}
\end{gather}
The wilkinoid shape is equal to the \(\beta \to 0\) limit of the cantoid, so
its planitude and alatude are given by:
\begin{gather}
  \label{eq:wilkinoid-Pi-Lambda}
  \text{wilkinoid} \quad
  \begin{cases}
    \quad \Pi &= \dfrac {5} {3}\\[10pt]
    \quad \Lambda &= \sqrt 3 \ .
  \end{cases}
\end{gather}
The wilkinoid results can also be obtained directly from
equation~\eqref{eq:wilkinoid-R-theta}, and in the case of \(\Lambda\) this
has already been noted by several authors \citep{Cox:2012a,
  Meyer:2016a}.

@@

The asymptotic opening angle
of the far wings, \(\theta_\infty\), can be found from
equation~\eqref{eq:ancantoid-theta-theta1-implicit} for the
ancantoids, together with the condition that
\(\theta_\infty + \theta_{1\infty} = \pi\).  These yield the implicit equation
\begin{equation}
  \label{eq:ancantoid-theta-inf}
  \theta_\infty - \left( \frac {k + 2 (1 - \beta)} {k + 2} \right) \tan \theta_\infty
  = \pi + 2 \beta I_k(\pi/2) \ ,
\end{equation}
where
\begin{equation}
  I_k(\pi/2) = \frac{\sqrt \pi}{4} \,
      \frac{  \GammaFunc\left( \frac{k+1}{2} \right)} {\GammaFunc\left(\frac{k+4}{2}\right)}
\end{equation}
and \(\GammaFunc\) is the usual Gamma function.  This can be compared with the equivalent result obtained by \CRW{} for the cantoids:
\begin{equation}
  \label{eq:cantoid-theta-inf}
  \theta_\infty - \tan \theta_\infty = \frac{\pi}{1 - \beta} \ .
\end{equation}
Note that, unlike in the cases of \(\Pi\) and \(\Lambda\),
equation~\eqref{eq:ancantoid-theta-inf} does \emph{not} reduce to
equation~\eqref{eq:cantoid-theta-inf} in the limit \(k \to 0\).  This is
because, for \(\theta > \ang{90}\), the \(k = 0\) ancantoid differs from the
cantoid since the former has no wind in the backward hemisphere (see
Figure~\ref{fig:anisotropic-arrows}).  Therefore there is less inner
support for the far wings of the bow, and so \(\theta_\infty\) is smaller than
in the cantoid case.  The wilkinoid result again follows from
\(\beta \to 0\), implying that \(\theta_\infty = \pi\), or, in other words, that the far
wings are asymptotically parallel to the symmetry axis, as is the case
for the paraboloid (App.~\ref{app:parabola}).  In the case of the
wilkinoid, however, the behavior is cubic in the wings,
\(z \sim r^3\), as opposed to quadratic as in the paraboloid.
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "quadrics-bowshock"
%%% End:

>>>>

\message{ !name(quadrics-bowshock.tex)}%\RequirePackage{amsmath}
\documentclass[useAMS, usenatbib, a4paper]{mnras}
\pdfsuppresswarningpagegroup=1

% The following is needed to fix the margins if using Letter-size paper
% REMOVE if your LaTeX uses A4 paper by default
%\addtolength\topmargin{-1.8cm}

% Standard LaTeX packages
% \usepackage[varg]{txfonts}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{color}
\usepackage{enumerate}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{hyperref}

\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}

% Define a fermata symbol for use inline with text
% 04 Jul 2017 WJH
% Based on info in Secs 18.4.1 and 24.1 of musixdoc.pdf
\usepackage{musixtex}
\makeatletter
\newcommand\textfermata{%
  {\let\extractline\relax
    \setlines10\smallmusicsize \nobarnumbers \nostartrule
    \staffbotmarg0pt \setclefsymbol1\empty \global\clef@skip0pt
    \raisebox{0ex}[0ex][0ex]{
      \startextract\addspace{-\afterruleskip}%
      \notes\fermataup{-2}\en
      \zendextract}}}
\makeatother

% % The \eye command is in the dingbat package, but we need to take care
% % of a naming conflict with the AMS package
% \usepackage{savesym}
% \savesymbol{checkmark}
% \usepackage{dingbat}

% The "eye" symbol is drawn with the \faEye command of the fontawesome
% package
\usepackage{fontawesome}
% The ``irregular'' symbol is drawn from the Icelandic Sorcery and Witchcraft 
\usepackage{staves}

% Fonts 
\usepackage{newtxtext}
% Note: newtxmath must come AFTER newtxtext
\usepackage[varvw,smallerops]{newtxmath}

\hypersetup{colorlinks=True, linkcolor=blue!50!black, citecolor=black,
  urlcolor=blue!50!black}

\usepackage{etoolbox}
\robustify\bfseries
\robustify\itshape

%% Bold italic
\newcommand\hmmax{0}            % we don't need heavy fonts
\newcommand\bmmax{1}            % reduce use of math alphabets for bold
\usepackage{bm}

%% Bundled custom packages
\usepackage{aastex-compat}

% Definitions needed in abstract
\newcommand\hii{\ion{H}{ii}}

\title[Bow shock shapes]{True versus apparent shapes of bow shocks and bow waves}

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Tarango Yong \& Henney]{
  Jorge A. Tarango Yong \& William J. Henney\\
  \AddressCRyA
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2017}

\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\Sin}{\mathcal{S}}
\DeclareMathOperator{\Cos}{\mathcal{C}}
\DeclareMathOperator{\Cot}{\mathcal{T}}
\DeclareMathOperator{\GammaFunc}{\Gamma}
\newcommand\w{\ensuremath{\mathrm{w}}}
\newcommand\C{\ensuremath{\mathrm{c}}}
\providecommand{\abs}[1]{\lvert#1\rvert}

\begin{document}

\message{ !name(app-shape-parameters.tex) !offset(-107) }

\section{Analytic derivation of thin-shell bow shapes}
\label{sec:thin-shell-shapes}

In this appendix, we provide analytic calculations of the planitude
(exact) and alatude (approximate) for the wilkinoid, cantoids, and
ancantoids.  We first consider the most general case of the
ancantoids, and then show how results for cantoids and the wilkinoid
follow as special cases.

@@

To find the alatude, \(\Lambda = R_{90} / R_0\), we use equation~\eqref{eq:crw-angles} at \(\theta = 90^\circ\) to write
\begin{equation}
  \label{eq:Lambda-from-theta-1-90}
  \Lambda = \frac{D} {R_0} \tan \theta_{1,90} \ , 
\end{equation}
where \(\theta_{1,90} = \theta_1(\theta = 90^\circ)\), which, following
equation~\eqref{eq:ancantoid-theta-theta1-implicit}, must satisfy
\begin{equation}
  \label{eq:theta-1-90-implicit}
  \theta_{1,90} \cot \theta_{1,90}  = 1 - \frac{2 \beta}{k + 2} \ . 
\end{equation}
Combining \eqref{eq:Lambda-from-theta-1-90} and
\eqref{eq:theta-1-90-implicit} with \eqref{eq:again-R0-over-D} yields
\begin{equation}
  \label{eq:Lambda-beta-xi-theta-1-90}
  \Lambda = \frac{ \left(1 + \beta^{1/2}\right) \,\theta_{1,90}} {\beta^{1/2} \left(1 - \xi_k \beta\right)} \ ,
\end{equation}
where
\begin{equation}
  \label{eq:xi-k}
  \xi_k = \frac{2} {k +2} \ .
\end{equation}
We now take the Taylor expansion of equation~\eqref{eq:theta-1-90-implicit} to find
\begin{equation}
  \label{eq:theta-1-90-Taylor}
  \theta_{1,90}^2 + \tfrac{1}{15}  \theta_{1,90}^4 + \mathcal{O}(\theta_{1,90}^6)
  = 3 \xi_k \beta \ , 
\end{equation}
which, if \(\theta_{1,90}\) is small, has the approximate solution
\begin{equation}
  \label{eq:theta-1-90-approx}
  \theta_{1,90} \approx \left( \frac{3 \xi_k \beta} {1 + \tfrac15 \xi_k \beta} \right) \ .
\end{equation}
Substituting back into equation~\eqref{eq:Lambda-from-theta-1-90}
yields an approximate value for the alatude of
\begin{equation}
  \label{eq:Lambda-approx}
  \text{ancantoid} \quad
  \Lambda \approx \frac {(3 \xi_k)^{1/2} \left( 1 + \beta^{1/2} \right)}
  { \left( 1 + \tfrac15 \xi_k \beta \right)^{1/2} \left( 1 - \xi_k \beta \right)} \ .
\end{equation}
This approximation is surprisingly accurate, with a relative error of
order \(1\%\) even for \(\beta\) as large as \(0.5\) with \(k = 0\).  The
leading term in the relative error scales as \((\beta / (k + 2))^2\), so the
approximation is even better for smaller \(\beta\) and larger \(k\).

@@

Since \(\Pi\) and \(\Lambda\) depend on only that portion of the inner wind
emitted in the forward hemisphere, \(\theta \le 90^\circ\), the results for the
cantoids can be found by taking \(k = 0\), in which case
equations~(\ref{eq:C-k-beta}, \ref{eq:final-planitude}, \ref{eq:xi-k},
\ref{eq:Lambda-approx}) yield
\begin{gather}
  \label{eq:cantoid-Pi-Lambda}
  \text{cantoid} \quad
  \begin{cases}
    \quad \Pi &= \dfrac {5} {3 \left( 1 - \beta^{1/2} \right)}\\[12pt]
    \quad \Lambda &= \dfrac { \sqrt 3} {\left( 1 + \tfrac15 \beta \right)^{1/2} \left( 1 - \beta^{1/2} \right)} \ .
  \end{cases}
\end{gather}
The wilkinoid shape is equal to the \(\beta \to 0\) limit of the cantoid, so
its planitude and alatude are given by:
\begin{gather}
  \label{eq:wilkinoid-Pi-Lambda}
  \text{wilkinoid} \quad
  \begin{cases}
    \quad \Pi &= \dfrac {5} {3}\\[10pt]
    \quad \Lambda &= \sqrt 3 \ .
  \end{cases}
\end{gather}

@@

The asymptotic opening angle
of the far wings, \(\theta_\infty\), can be found from
equation~\eqref{eq:ancantoid-theta-theta1-implicit} for the
ancantoids, together with the condition that
\(\theta_\infty + \theta_{1\infty} = \pi\).  These yield the implicit equation
\begin{equation}
  \label{eq:anacantoid-theta-inf}
  \theta_\infty - \left( \frac {k + 2 (1 - \beta)} {k + 2} \right) \tan \theta_\infty
  = \pi + 2 \beta I_k(\pi/2) \ ,
\end{equation}
where
\begin{equation}
  I_k(\pi/2) = \frac{\sqrt \pi}{4} \,
      \frac{  \GammaFunc\left( \frac{k+1}{2} \right)} {\GammaFunc\left(\frac{k+4}{2}\right)}
\end{equation}
and \(\GammaFunc\) is the usual Gamma function.  This can be compared with the equivalent result obtained by \CRW{} for the cantoids:
\begin{equation}
  \label{eq:cantoid-theta-inf}
  \theta_\infty - \tan \theta_\infty = \frac{\pi}{1 - \beta} \ .
\end{equation}
Note that, unlike the case for \(\Pi\) and \(\Lambda\),
equation~\eqref{eq:anacantoid-theta-inf} does \emph{not} reduce to
equation~\eqref{eq:cantoid-theta-inf} in the limit \(k \to 0\).  This is
because, for \(\theta > 90^\circ\), the \(k = 0\) ancantoid differs from the
cantoid since there is no wind in the backward hemisphere (see
Figure~\ref{fig:anisotropic-arrows}).  Therefore there is less inner
support for the far wings of the bow, and so \(\theta_\infty\) is smaller than
in the cantoid case.  The wilkinoid result again follows from
\(\beta \to 0\), implying that \(\theta_\infty = \pi\), or, in other words, that
the far wings are asymptotically parallel to the symmetry axis, as is
the case for the paraboloid (App.~\ref{app:parabola}).  In the case of
the wilkinoid, however, the behavior is cubic in the wings,
\(z \sim r^3\), as opposed to quadratic as in the paraboloid.
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "quadrics-bowshock"
%%% End:

\message{ !name(quadrics-bowshock.tex) !offset(-137) }

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2017-12-06 papers/Papers1234/app-r90.tex
Submitted first paper! To MNRAS and arxiv

\appendix
% \appendixpage
% \addappheadtotoc
\section{Parabola of Revolution}
\label{app:parabola}

In this section we develop the particular case of the parabola.
The parametrization is given by:

\begin{align}
x &= - \frac{1}{2}R_ct^2 + R_0 \\
y &= R_c t
\end{align}
Where $R_c$ is the radius of curvature and $R_0$ is the distance of the parabola nose to the
origin.
The respective tridimensional shape is given by:
\begin{align}
x &= -\frac{1}{2}R_ct^2 + R_0 \label{eq:x-par-a}\\
y &= R_c t \cos\phi  \label{eq:y-par-a}\\
z &= R_c t \sin\phi  \label{eq:z-par-a}
\end{align}
The azimutal angle where the surface is tangent to the line of sight in this case is given by:
\begin{align}
\sin\phi_t = -\frac{\tan i}{t} \label{eq:sin-tan-a} 
\end{align}

Subtituting (\ref{eq:sin-tan-a}) into (\ref{eq:x-par-a}), (\ref{eq:y-par-a}) and
(\ref{eq:z-par-a}) we find the apparent shape
of the paraboloid:

\begin{align}
x' &= -\frac{R_c(\frac{1}{2}t^2 \cos^2 i -\sin^2 i)}{\cos i}+R_0\cos i \\
y' &= R_c\left(t^2-\tan^2 i\right)^{1/2} 
\end{align}

Taking the limit of equations (\ref{eq:qprime}) and (\ref{eq:Aprime}) when $\theta_c$ tends to zero we find that:

\begin{align}
\left(\frac{q'}{q}\right)_{\mathrm{parabola}} &= 1+\frac{\tilde{R}_c\tan^2 i}{2}\\
\left(\tilde{R}'_c\right)_{\mathrm{parabola}} &= \frac{\tilde{R}_c}{\cos^2 i + \frac{\tilde{R}_c}{2}\sin^2 i}
\end{align}

\section{Analytic derivation of the radius of curvature in the Thin Shell model}
\label{app:rc-analytic}

For small $\theta$ we may do a polynomial expansion for the shell shape such as:

\begin{align}
R \simeq R_0\left(1+\gamma\theta^2 + \Gamma\theta^4\right) \label{eq:R-exp}
\end{align}

The radius of curvature at the axis for $R$ is given by:

\begin{align}
R_c = R_0\left(1-2\gamma\right)^{-1}
\end{align}

The coefficient gamma may be derived by an expansion at small angles of equation
(\ref{eq:th1th}), as follows:

From the first term of the right side we get:

\begin{align}
\cot\theta &\simeq \theta^{-1}\left[1-\frac{1}{3}\theta^2\right] \\
\cos^k\theta\sin^2\theta &\simeq \theta^2 - \left(\frac{1}{3} + \frac{k}{2}\right)\theta^4 \\
\implies I_k(\theta) &\simeq \frac{1}{3}\theta^3\left[ 1 - \frac{1}{10}(3k+2)\theta^2\right]\\
\implies 2\beta I_k(\theta)\cot\theta &\simeq \frac{2}{3}\beta\theta^2\left[1-\frac{1}{30}
(9k+16)\theta^2\right]\label{eq:AR1} 
\end{align}

For the second term we get:

\begin{align}
-\frac{2\beta}{k+2}\left(1-\cos^{k+2}\theta\right) & \simeq -\beta\theta^2\left[1-\frac{1}{12}
(3k+4)\theta^2\right] \label{eq:AR2}
\end{align}

For the left side we use equation (25) from \CRW{}. Then, equation (\ref{eq:th1th}) results
as follows:

\begin{align}
\theta_1^2\left(1+\frac{1}{15}\theta_1^2\right) = \beta\theta^2\left[1+\frac{1}{60}(4-9k)
\theta^2\right] \label{eq:th1th-app}
\end{align}

And we can use the approximation $\theta_1 \approx \beta\theta^2$ for the correction term in
the left side of (\ref{eq:th1th-app}):

\begin{align}
\theta_1^2 &= \beta\theta^2\left[1+\frac{1}{60}(4-9k)\theta^2\right]
\left(1-\frac{\beta}{15}\theta^2\right) \\
\implies \theta_1^2 &= \beta\theta^2\left[1+ 2C_{k\beta}\theta^2\right]
\label{eq:th1th-small}\\
\mathrm{where:~} C_{k\beta} &\equiv \frac{1}{2}\left(A_k-\frac{\beta}{15}\right) \\
A_k &\equiv \frac{1}{15}-\frac{3k}{20}
\end{align}

Now, using equation (23) from \CRW{} we may estimate $R$ at low angles. To do this, we need to
expand each term as follows (neglecting terms of order four or higher):

\begin{align}
\theta_1 = &= \beta^{1/2}\theta\left[1+ 2C_{k\beta}\theta^2\right]^{1/2} \\
\theta + \theta_1 &= \theta\left[1+\beta^{1/2}\left(1+2C_{k\beta}\theta^2\right)\right]\\
\sin\theta_1 &= \theta_1\left[1-\frac{\theta_1^2}{6}\right] \\
 &= \beta^{1/2}\theta\left[1+\left(C_{k\beta}-\frac{1}{6}\beta\right)\theta^2\right] \\
 \sin(\theta+\theta_1) &= \left[\theta+\theta_1\right]\left[1-\frac{\left(\theta+\theta_1
 \right)^2}{6}\right] \\
 &= \theta\left(1+\beta^{1/2}\right)\left\lbrace 1+\left[\frac{C_{k\beta}\beta^{1/2}}
 {1+\beta^{1/2}}-\frac{1}{6}\left(1+\beta^{1/2}\right)^2\right]\theta^2\right\rbrace
\end{align}

So, combining these terms with equation (23) from \CRW{} we found the final expression for $R$:

\begin{align}
\frac{R}{D}\equiv \frac{\sin\theta_1}{\sin(\theta+\theta_1)} = \frac{\beta^1/2}{1+\beta^{1/2}}
\left\lbrace 1 + \theta^2\left[\frac{C_{k\beta}}{1+\beta^{1/2}}+\frac{1}{6}\left(1+2\beta^{1/2}
\right)\right] \right\rbrace \label{eq:r-small-theta}
\end{align}

Returning to equation (\ref{eq:R-exp}) we see the following:

\begin{align}
R_0 &= \frac{\beta^{1/2}}{1+\beta^{1/2}} \\
\gamma &= \frac{C_{k\beta}}{1+\beta^{1/2}}+\frac{1}{6}\left(1+2\beta^{1/2}\right)
\label{eq:app-gamma}
\end{align}

We recover equation (27) of \CRW{} for $R_0$ and equation (\ref{eq:app-gamma}) is the
needed term to calculate the radius of curvature at the axis.

@@

\section{Derivation of Characteristic Radii in Isotropic Wind/Parallel interaction Problem}
\label{app:ch-rad-Wilkin}

$\tilde{R}_{90}$ is obtained by simply evaluating equation (\ref{eq:R-Wilkin}) at $\theta=\frac{\pi}{2}$.
For the Radius of curvature we follow a similar procedure than the Wind/Wind interaction, but using equation
(\ref{eq:R-Wilkin}) for $R(\theta)$ and inserting the cosecant into the square root.

Expanding the terms of $R(\theta)$ we find the following:

\begin{align}
  \csc^2\theta &\simeq \theta^{-2}\left[1+\frac{\theta^2}{3}\left(1+\frac{\theta^2}{5}\right)\right] \\
  1-\theta\cot\theta &\simeq \frac{\theta^2}{3}\left[1 + \frac{\theta^2}{15}\left(1+\frac{2\theta^2}{21}\right)\right] \\
  \implies \tilde{R}(\theta) \simeq 1 + \frac{\theta^2}{5} + O\left[\theta^4\right]
\end{align}

From equation (\ref{eq:Rcurv}) for the radius of curvature we finally get the numerical value for $\tilde{R_c} = \frac{5}{3}$
\clearpage

% Landscape table needs to go on its own page, so we need to manually
% adjust the section counter so that it will ``belong'' to the next
% section
\addtocounter{section}{1}
\sisetup{detect-all=true, detect-inline-weight=math}
\sisetup{round-mode=figures, round-precision=2}
\begin{landscape}
\begin{table}
  \setlength\tabcolsep{3pt}
  \caption{Results of all statistical tests performed on observed bow shock shape parameters. Significant correlations are shown in \textbf{bold}, marginally significant correlations in \textit{italic}}
  \label{tab:big-p}
  \input{figs/mipsgal-summary-stats-table-body.tex}
\end{table}
\end{landscape}
\begin{table}
  \contcaption{Results of all statistical tests performed on observed
    bow shock shape parameters.}
  \begin{tabular}{p{0.9\linewidth}}
    \toprule
    \textit{Description of columns:}
    (Col.~1)~How the two A/B source sub-samples are defined, also giving the size of each sub-sample, \(n_{\text{A}}\) and~\(n_{\text{B}}\).
    (Col.~2)~Dependent variable whose distribution is compared between the two sub-samples.
    (Cols.~3--6)~Mean and standard deviation, \(\sigma\), of the dependent variable for each of the two sub-samples.
    (Cols.~7--8)~Mean over each sub-sample of the observational dispersion (\(\epsilon\), standard deviation) of radii that contribute to the dependent variable for each individual source, as in steps~\ref{step:R0} and \ref{step:R90} of \S~\ref{sec:autom-trac-fitt}.  Note that in the case of \(R_c\), this is \(\epsilon(R_0)\), and so is not a direct measure of the observational uncertainty in \(R_c\). 
    (Cols.~9--10)~``Standard error of the mean'' (s.e.m.) of the dependent variable for each of the two sub-samples. 
    (Cols.~11--13)~Standardized ``effect sizes'', which are dimensionless measures of the difference in the distribution of the dependent variable between the two sub-samples.
    (Col.~11)~Rank biserial correlation coefficient \citep{Cureton:1956a}, which is obtained by considering all \(n_{\text{A}} n_{\text{B}}\) pair-wise comparisons of the dependent variable between a source in sub-sample~A and a source in sub-sample~B.  It is the difference between the fraction of such comparisons ``won'' by sub-sample~A and those ``won'' by sub-sample~B, and thus may vary between \(-1\) and \(+1\). 
    (Col.~12)~Cohen's \(d\), which is a dimensionless mean difference: \(d = (\langle \text{A} \rangle - \langle \text{B} \rangle) / \sigma_{\text{pool}} \), where \(\sigma_{\text{pool}} = (n_{\text{A}} \sigma_{\text{A}}^2 + n_{\text{B}} \sigma_{\text{B}}^2)^{1/2} / \sqrt{n_{\text{A}} + n_{\text{B}}}\) is the pooled standard deviation.
    (Col.~13)~Ratio of standard deviations between the two sub-samples.
    (Cols.~14--16)~Probabilities (\(p\)-values) of the two sub-samples being as different as observed if they were to be drawn from the same population, according to three different non-parametric tests.
    (Col.~14)~Anderson--Darling 2-sample test, which is a general test of similarity between two distributions that is designed to retain sensitivity to differences in the tails of the distributions.
    (Col.~15)~Mann--Whitney--Wilcoxon \(U\) test \citep{Mann:1947a}, which is sensitive 
    (Col.~16)~Brown--Forsythe test for equality of variance \citep{Brown:1974a}
    \\
    \bottomrule
  \end{tabular}
\end{table}

\addtocounter{section}{-1}

\section{Distribution of p-values for all correlations tested}
\label{sec:distr-p-values}
\addtocounter{table}{1}

\begin{figure}
  (a)\\
  \includegraphics[width=\linewidth]{figs/p-value-histogram-new-linear}\\
  (b)\\
  \includegraphics[width=\linewidth]{figs/p-value-histogram-new}
  \caption{Histogram of \(p\)-values for all non-parametric 2-sample
    tests listed in Table~\ref{tab:big-p}. (a)~Uniformly spaced linear
    bins and linear vertical axis. (b)~Uniformly spaced logarithmic
    bins and logarithmic vertical axis, with all values
    \(p \le 10^{-6}\) included in the leftmost bin.  Short thin vertical
    lines above the horizontal axis show the individual value.  The
    thick vertical dashed lines shows the traditional threshold values
    for significance: \(p = 0.003\) (\(\approx 3 \sigma\)) and
    \(p = 0.05\) (\(\approx 2 \sigma\)). The red solid line shows the uniform
    distribution of \(p\)-values that would be expected if the null
    hypothesis were always true, that is, if no significant
    correlations existed.}
  \label{fig:histo-p-values}
\end{figure}

Results from all the statistical tests discussed in \S~\ref{sec:corr-shape} are given in Table~\ref{tab:big-p}.

\begin{table}
  \caption{Lower bounds on the Type~I error rate, \(\alpha\)}
  \label{tab:type-I}
  \centering
  \begin{tabular}{lrr} \toprule
    Correlation & \(p\) & \(\alpha\) \\
    \midrule
    \multicolumn{3}{l}{\itshape Quantitative \dotfill}\\
    \(R_c\) vs \(R_0\) & \textit{0.0054} & \textit{0.0712} \\
    \(R_{90}\) vs \(R_0\) & \textbf{0.0001} & \textbf{0.0025} \\
    \(R_c\) vs \(H_0\) & 0.1229 & 0.4119 \\
    \(R_{90}\) vs \(H_0\) & \textit{0.0215} & \textit{0.1833} \\
    \(R_c\) vs \(A_K\) & 0.19 & 0.4617 \\
    \(R_{90}\) vs \(A_K\) & 0.63 & 1 \\
    \(R_c\) vs \(|b|\) & 0.19 & 0.4617 \\
    \(R_{90}\) vs \(|b|\) & 0.31 & 0.4967 \\
    \(R_c\) vs \(\cos(\ell)\) & 1.00 & 1 \\
    \(R_{90}\) vs \(\cos(\ell)\) & 0.36 & 0.4999 \\
    \midrule
    \multicolumn{3}{l}{\itshape Categorical} \dotfill\\
    \(R_c\) vs Facing & 0.71 & 1 \\
    \(R_c\) vs H II & 0.50 & 1 \\
    \(R_{90}\) vs Facing & 0.60 & 1 \\
    \(R_{90}\) vs H II & 0.52 & 1 \\
    \(R_c\) vs Multiple & 1.00 & 1 \\
    \(R_{90}\) vs Multiple & 0.34 & 0.4993 \\
    \(R_c\) vs \SI{8}{\um} & 0.82 & 1 \\
    \(R_{90}\) vs \SI{8}{\um} & 0.6 & 1 \\
    \midrule
    \multicolumn{3}{l}{\itshape Other datasets} \dotfill\\
    \(R_c\) vs Herschel & 0.074 & 0.3437 \\
    \(R_{90}\) vs Herschel & \textbf{0.00052} & \textbf{0.0106} \\
    \(R_c\) vs M42 & \textbf{0.00048} & \textbf{0.0099} \\
    \(R_{90}\) vs M42 & \textbf{0.0000105} & \textbf{0.0003} \\
    \bottomrule
  \end{tabular}
\end{table}

The \(p\)-values are the probability of finding a difference between
two populations as large as (or larger than) what is observed,
\emph{given} that there is no difference in the underlying
distribution from which the two populations are drawn (that is, given
that the null hypothesis is true).  However, what we really want to
know is something else: the probability that the null hypothesis is
true, \emph{given} the observations.  That is, the probability,
\(\alpha\), of a \textit{false positive}, also known as the \textit{Type I
  error rate}.  The common mistake of conflating these two definitions
is known as the ``\(p\)-value fallacy'' \citep{Goodman:1999a}, or``the
error of the transposed conditional'', as discussed in detail by
\citet{Colquhoun:2014a}.  It is possible to derive \(\alpha\) from
\(p\) using Bayes' theorem (e.g., \citealp{Goodman:1999b}), but that
requires an estimate of the prior probability of the null hypothesis,
independent of the observations.  Alternatively, it is also possible
to find a lower bound on \(\alpha\) from a frequentist approach
\citep{Sellke:2001a}:
\begin{equation}
  \label{eq:type-I}
  \alpha(p) \ge \bigg[ 1 - \big(e\, p \ln p\big)^{-1} \bigg]^{-1}
  \quad \text{valid for } p < 1/e.
\end{equation}
This is the approach we adopt here, which also numerically coincides
with the Bayesian approach for the case where the prior probability of
the null hypothesis is 0.5.  The reason that this is only a lower
limit for \(\alpha\) is that if we have overwhelming a priori evidence that
the null hypothesis is true (for instance, from previous empirical
studies, or because it follows from a well-supported theory), then a
Bayesian calculation would give a much higher value of \(\alpha\) than
\eqref{eq:type-I} does.  In our case, however, we have no strong
reasons for favoring any of the null hypotheses, so it is reasonable
to assume \(\alpha\) is close to the lower limit given in \eqref{eq:type-I}.

In order to choose a threshold \(p\)-value that counts as a
``significant'' result, one then needs to balance the risks of false
positives against the risks of \textit{false negatives}.  The false
negative probability, \(\beta\), also known as \textit{Type II error
  rate}, is the probability of failing to reject an untrue null
hypothesis.  That is, in the context of this paper, it is the
probability of failing to detect a real difference between two
sub-samples, or a real correlation between two variables.  The
complementary probability, \(1 - \beta\), is known as the
\textit{statistical power} or sensitivity of the test.  The value of
\(\beta\) depends on three factors:
\begin{enumerate}[1.]
\item The \textit{effect size}, which is a measure of the magnitude of
  the difference in a dependent variable between two sub-samples, or
  the degree of correlation between two continuous variables.  For the
  two sub-sample case, it is common to use a standardised mean
  difference, such as Cohen's \(d\) statistic \citep{Cohen:1988a}:
  \(d = (\bar{X}_A - \bar{X}_B) / s\), where \(\bar{X}_A\),
  \(\bar{X}_B\) are the means of the dependent variable \(X\) for
  samples A and B, while \(s\) is the pooled standard deviation of
  \(X\).  For the case of two continuous variables, the Pearson linear
  correlation coefficient, \(r\), can be used.  In both cases, rules
  of thumb have been developed \citep{Ruscio:2008a} for classifying an
  effect as ``large'' (\(d > 0.8\), \(r > 0.4\)) or ``small''
  (\(d < 0.2\), \(r < 0.1\)).  Alternatively, non-parametric
  statistics can be used, such as the \(A\) measure of stochastic
  superiority \citep{Delaney:2002a}.
\end{enumerate}

Obviously, this depends on the \textit{effect size}, which is the

All astronomical data analysis is \emph{post hoc} analysis, since the universe was not set up to test a particular hypothesis (as far as we know).  It is therefore important to guard against the ``multiple comparisons problem'', whereby seemingly significant correlations are found where none really exist, simply by virtue of the large number of tests that were carried out.

Under the more conservative Holm--Bonferroni method, only comparisons with \(p < 0.001\) would be significant.

The p-curve \citep{Head:2015a}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "quadrics-bowshock.tex"
%%% End:

>>>>

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "quadrics-bowshock"
%%% End:
